---
title: "School Data Time Series"
author: "Rob Wells. Thanks to Mary Hennigan, Paige McGaughy, Abby Zimmardi, Katy Seiter" 
date: "9/12/2021"
output: pdf_document
---

# Compilation of School COVID-19 Reports for Arkansascovid.com Calculations

#Instructions on how to prepare the spreadsheet for importing
https://docs.google.com/document/d/1meNe49P1cJGJut4GaTHI2qo5UJ7PMpRGkkqVBm5E1LQ/edit?usp=sharing


#Here's a video on how to process PDFs into data
https://video.uark.edu/media/Capturing+Data+from+ADH+School+Reports+and+Cleaning+In+R/1_1il3ruhu

#Retrieving data from biweekly ADH school report: Educational Institutions:
#https://www.healthy.arkansas.gov/programs-services/topics/covid-19-reports

- **This notebook retrieves data from an Excel sheet and compiles a time series of ADH school infection reports for the Fall 2021 school year**

-**It creates a masterfile: schools_master_2021.csv**

# Part 1: Import Data, Clean It

```{r include=FALSE}
#install.packages("slider")
#install.packages("zoo")
#install.packages("gtools")
library(tidyverse)
library(janitor)
library(jsonlite)
library(gtools)
library(zoo)
library(reshape2)
library(slider)
```

#Type in the report date of the ADH document
```{r}
reportdate <- "2021-09-09"
```

#Import the Excel sheet
```{r include=FALSE}

schoolraw <- rio::import("/Users/robwells/Dropbox/Classes/Data Fall 2020/ArkansasCovid/Schools/School_Report_09092021.xlsx")
schoolraw <- janitor::clean_names(schoolraw)

#name columns
schoolraw <- schoolraw %>% 
   rename(name = timeframe_8_1_2021_present, cumulative_active_cases = x4, cumulative_faculty_staff_cases2 = x6, cumulative_student_cases2 = x8, cumulative_total3 = x10)

#add date field
schoolraw$Date <- Sys.Date()
#if you are doing this a day later, use this instead. -1 is yesterday. -2 is two days etc
#schoolraw$Date <- Sys.Date()-1

schoolraw <- schoolraw %>% 
   select(Date, type, name, active_cases, cumulative_active_cases, cumulative_faculty_staff_casesa, cumulative_student_casesa, cumulative_totalb, cumulative_faculty_staff_cases2, cumulative_student_cases2, cumulative_total3)
   
schoolraw <- schoolraw %>% 
   rename(Type = type, Name = name, Active_Cases = active_cases, Cumulative_Faculty_Staff_Cases = cumulative_faculty_staff_casesa, Cumulative_Student_Cases = cumulative_student_casesa, Cumulative_Total=cumulative_totalb, Cumulative_Active_cases=cumulative_active_cases, All_Faculty_Staff_Total = cumulative_faculty_staff_cases2, All_Student_Total = cumulative_student_cases2, Cumulative_Group_Total = cumulative_total3)


#Change this date to reflect the report date, ie Sept 9 etc you're working on
schoolraw <- schoolraw[!grepl("cSeptember 06, 2021", schoolraw$Name),]

#
schoolraw <- schoolraw[!grepl("Timeframe", schoolraw$Name),]
schoolraw <- schoolraw[!grepl("Counts of less than", schoolraw$Name),]
schoolraw <- schoolraw[!grepl("cases may not add up to", schoolraw$Name),]
schoolraw <- schoolraw[!grepl("Counts of less than five", schoolraw$Name),]
schoolraw <- schoolraw[!grepl("totals", schoolraw$Name),]
schoolraw <- schoolraw[!grepl("Cumulative", schoolraw$All_Faculty_Staff_Total),]

schoolraw[4:11] <- lapply(schoolraw[4:11], as.numeric)


schoolname <- paste("schools_", reportdate,".csv",sep="")
write.csv(schoolraw, file = schoolname)

```





#Add to master file
```{r}
schools_master_2021 <- rio::import("https://raw.githubusercontent.com/Arkansascovid/Main/master/Schools/schools_master_2021.csv")

#from hard drive
# schools_master_2021 <- rio::import("/Users/robwells/Dropbox/Classes/Data Fall 2020/ArkansasCovid/CovidFall2020/schools_master_2021.csv")

schools_master_2021 <- schools_master_2021[ -c(1) ]

schools_master_2021 <- rbind(schoolraw, schools_master_2021)

write.csv(schools_master_2021, "Schools/schools_master_2021.csv")
```

#This part carves off the totals into a separate table and carves off specific schools into a specific table
```{r}
school_master_totals <- schools_master_2021 %>% 
   filter(Type=="Totals for all schools.") %>% 
   select(Date, Name, Cumulative_Active_cases, All_Faculty_Staff_Total, All_Student_Total, Cumulative_Group_Total) %>% 
   arrange(desc(Name)) %>% 
   mutate(New_Active_Cases = (Cumulative_Active_cases-lead(Cumulative_Active_cases))) %>% 
   mutate(Pct_Chg_Active = (Cumulative_Active_cases-lead(Cumulative_Active_cases))/lead(Cumulative_Active_cases)) 


school_master_totals$Pct_Chg_Active <- formattable::percent(school_master_totals$Pct_Chg_Active)


x <- school_master_totals %>% 
   filter(Date=="2021-08-16") %>% 
   mutate(Pct_Chg_Active = "-") %>% 
   mutate(New_Active_Cases = "-")

y <- school_master_totals %>% 
   filter(Date>"2021-08-16")

z <- smartbind(y,x)

z <- z %>% 
   arrange(desc(Name))

school_master_totals <- z

write.csv(school_master_totals, "Schools/school_master_totals.csv")
#Feeds this datawrapper table
#https://app.datawrapper.de/table/cTtxA/upload
```


#This part subsets just the named schools and cuts out the totals
```{r}
schools_details <- schools_master_2021 %>% 
   filter(!(Type=="Totals for all schools.")) %>% 
   filter(!(Type=="Totals among schools with less than 5 active cases.")) %>% 
   filter(!(Name=="TOTAL")) %>% 
   filter(!(Name=="PUBLIC SCHOOL DISTRICT")) %>% 
   filter(!(Name=="COLLEGE/UNIVERSITY")) %>% 
   filter(!(Name=="PRIVATE SCHOOL")) %>% 
   select(Date, Name, Cumulative_Total, Active_Cases, Cumulative_Student_Cases, Cumulative_Faculty_Staff_Cases, Type) %>% 
   arrange((Name))

schools_details

write.csv(schools_details, "Schools/school_details.csv")

#Feeds this datawrapper table
# https://app.datawrapper.de/table/5hdkd/upload


```
#To Run Schools Combo Separately, import data from here first
```{r}
# school <- rio::import("https://raw.githubusercontent.com/profrobwells/CovidFall2020/master/Schools/school_details.csv")

```


#This part calculates COVID per student population by joining the ADH school COVID report with Dept of Education school demographic report

```{r}
school <- schools_master_2021
school <- janitor::clean_names(school)
school$district <- tolower(school$name)
school$district <- gsub(" ", "_", school$district)
school$district <- gsub("southside_school_dist(independence)", "southside_school_dist_(independence)", school$district)
school[4:7] <- lapply(school[4:7], as.numeric)
school$date <- as.Date(school$date)
head(school)
```

```{r}
schoolethnicity <- rio::import('https://raw.githubusercontent.com/Arkansascovid/Main/master/Schools/School_Districts_FY31_Districts_EnrollmentByRace.csv')
schoolethnicity<- janitor::clean_names(schoolethnicity)
schoolethnicity$district <- tolower(schoolethnicity$district_description)
schoolethnicity$district <- gsub("district", "dist", schoolethnicity$district)
schoolethnicity$district <- gsub(" ", "_", schoolethnicity$district)
head(schoolethnicity)
```

#join tables
```{r}
schools_combo <- school %>% 
  inner_join(schoolethnicity, by=c("district"))
glimpse(schools_combo)
```


#Math
```{r}
schools_combo <- schools_combo %>% 
  mutate(Pct_Hispanic = (hispanic_total/student_total)) %>% 
   mutate(Pct_White = (white_total/student_total)) %>% 
   mutate(Pct_Black = (black_total/student_total))

factcheck <- schools_combo %>% 
  select(district,  student_total, hispanic_total, Pct_Hispanic, white_total, Pct_White, black_total, Pct_Black)


schools_combo$Pct_Hispanic <- round(schools_combo$Pct_Hispanic, 2)
schools_combo$Pct_White <- round(schools_combo$Pct_White, 2)
schools_combo$Pct_Black <- round(schools_combo$Pct_Black, 2)

factcheck
```

#Covid per 1000 students
```{r}
schools_combo <- schools_combo %>% 
  mutate(Student_Covid_Per1000=(cumulative_student_cases/student_total)*1000) %>% 
  mutate(Active_Covid_Per1000=(active_cases/student_total)*1000)
#Active_Covid_Per1000 could be an issue if the active cases are both students and faculty

schools_combo$Student_Covid_Per1000 <- round(schools_combo$Student_Covid_Per1000, 2)
schools_combo$Active_Covid_Per1000 <- round(schools_combo$Active_Covid_Per1000, 2)
```

```{r}
write.csv(schools_combo, "Schools/public_schools_combo.csv")
```

#College Enrollment and Cases

#Import Enrollment Data For Colleges
```{r}
collegeenroll <- rio::import("https://github.com/Arkansascovid/Main/blob/master/Schools/School%20College%20Enrollment%20Data.xlsx?raw=true", which = "NCES_College_info")

collegeenroll <- janitor::clean_names(collegeenroll)

collegeenroll$college <- gsub("CENTRAL BAPRIST COLLEGE, CONWAY", "CENTRAL BAPTIST COLLEGE, CONWAY", collegeenroll$college)

```

#join tables
```{r}
college_combo <- school %>% 
  inner_join(collegeenroll, by=c("name"="college")) 
glimpse(college_combo)

```

#Covid per 1000 college students
```{r}
college_combo <- college_combo  %>% 
  mutate(Student_Covid_Per1000=(cumulative_student_cases/student_population)*1000) %>% 
  mutate(Active_Covid_Per1000=(active_cases/student_population)*1000)
#Active_Covid_Per1000 could be an issue if the active cases are both students and faculty

college_combo$Student_Covid_Per1000 <- round(college_combo$Student_Covid_Per1000, 2)
college_combo$Active_Covid_Per1000 <- round(college_combo$Active_Covid_Per1000, 2)

write.csv(college_combo, "Schools/college_combo.csv")

```


This script feeds the following graphics:
https://app.flourish.studio/story/981386/edit

Searchable Table: Public Schools Active Case Rate
https://app.flourish.studio/visualisation/7222023/edit
Data:
https://raw.githubusercontent.com/Arkansascovid/Main/master/Schools/public_schools_combo.csv

Total Active COVID-19 Cases in Arkansas Educational Institutions
https://app.flourish.studio/visualisation/7222039/edit
Data: https://raw.githubusercontent.com/Arkansascovid/Main/master/Schools/school_master_totals.csv

Active COVID-19 Case Rate in Arkansas Public Schools
https://app.flourish.studio/visualisation/7222020/edit
Data
https://raw.githubusercontent.com/Arkansascovid/Main/master/Schools/public_schools_combo.csv

Searchable Table: College Active Case Rate
https://app.flourish.studio/visualisation/7222022/edit
Data:
https://raw.githubusercontent.com/Arkansascovid/Main/master/Schools/college_combo.csv

Active COVID-19 Case Rate in Arkansas Colleges
https://app.flourish.studio/visualisation/7222018/edit
Data:
https://raw.githubusercontent.com/Arkansascovid/Main/master/Schools/college_combo.csv

Active Case Rate by Date
https://app.flourish.studio/visualisation/7222019/edit
Data:
https://raw.githubusercontent.com/Arkansascovid/Main/master/Schools/college_combo.csv


